---
title: "GLM: mixed effects logisitc regression."
author: "Jacob Westaway"
date: "Last updated on `r Sys.Date()`"
output: pdf_document
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages.

```{r, warning=F, message=F, results='hide'}
sapply(c("lme4", "aods3", "tidyverse", "ggplot2", "MuMIn", "gridExtra", "effects", 
         "car", "emmeans", "ggpubr", "DESeq2", "phyloseq", "sjPlot", "grid"), 
         require, character.only = TRUE)
```


## Use DESeq2 to determine what taxa to include in the model.

### Define function for calculating geometric means.

```{r, warning=F, message=F, results='hide'}
calc_geo_means <- function(deseq_object){
# geometric mean
  gm_mean = function(x, na.rm = TRUE){
    exp(sum(log(x[x > 0]), na.rm = na.rm) / length(x))
  }
  geoMeans <- apply(counts(deseq_object), 1, gm_mean)
# size factors
  estimateSizeFactors(deseq_object, geoMeans = geoMeans) 
}
```

### Define function to subset out taxa with small counts and low occurance (count of at least **10** in **60** or more samples).

```{r, warning=F, message=F, results='hide'}
deseq_filter <- function(deseq_object){
  nc <- counts(deseq_object, normalized = TRUE)
  filtered <- rowSums(nc >= 10) >= 60 # filter = abundance of 10 in 60 samples.
  deseq_object[filtered,]
}
```

### Define function to extract significant results from a DESeq2 LRT test

```{r, warning=F, message=F, results='hide'}
get_deseq_res_lrt <- function(deseq_object){
  res = results(deseq_object)
  res = res[order(res$padj, na.last = NA), ]
  sigtab = res[(res$padj < 0.01), ] 
  sigtab = cbind(as(sigtab, "data.frame"), 
          as(tax_table(ps3.Microbiome)[rownames(sigtab), ], "matrix"))
  sigtab %>%
  arrange(padj) %>%
  select("log2FoldChange", "lfcSE", "padj", "Genus") 
}
```

### Run the DESeq2 analysis

```{r, message=F, warning=F}
phyloseq_to_deseq2(ps3.Microbiome, ~ parasite_burden) %>% 
    calc_geo_means() %>% 
    deseq_filter() %>% 
    DESeq(fitType = "local", test = "LRT", reduced = ~ 1) %>% 
    get_deseq_res_lrt() %>% 
    remove_rownames()
```


## Create a new dataframe that includes the transformed abundances of the significant genera above.
 
```{r, warning=F, message=F, results='hide'}
metadata_with_taxa <- subset_samples(ps3, Type == "Microbiome") %>% 
  sample_data() %>% 
  unclass() %>% 
  as.data.frame() %>% 
  mutate(ID = paste0("AM", ID)) %>% 
  left_join(
    (phyloseq_to_deseq2(ps3.Microbiome, ~ parasite_burden) %>% 
    calc_geo_means() %>% 
    counts(normalized = TRUE) %>%  
    as.data.frame() %>% 
    filter(rownames(.) == "TTTCGAATCATTCACAATGGGGGAAACCCTGATGGTGCAACGCCGCGTGGGGGATGAAGGCCTTCGGGTTGTAAACCTCTGTCACCAAGGAGCAACAAGCCGGTTCATAGCCGGCCCTGAGTTAACTTGGAGAGGAAGCAGTGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGACTGCAAGCGTTACTCGGATTCACTGGGCGTAAAGGGTGCGTAGGCCGCTAAGCGTGTCGGGTGTGAAATCTCGGGGCTCAACCTCGAAACTGCGCCCGAAACTGTTTAGCTAGAGTGTCGGAGAGGTAAGCGGAATTCCAGGTGTAGCGGTGAAATGCGTAGATATCTGGAGGAACACCAATGGCGAAGGCAGCTTACTGGACGACAACTGACGCTGAGGCACGAAAGCGTGGGTAGCGAAAGGG" | 
             rownames(.) ==  "TGAGGAATATTGGACAATGGACGAAAGTCTGATCCAGCCATGCCGCGTGCAGGATGACGGCCCTATGGGTTGTAAACTGCTTTTATACAGGAAGAAACACTCCCACGTGTGGGGGCTTGACGGTACTGTACGAATAAGGATCGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGGATCCAAGCGTTATCCGGAATTATTGGGTTTAAAGGGTCCGCAGGCGGTCTATTAAGTCAGAGGTGAAATCTTGCAGCTCAACTGTAAAATTGCCTTTGATACTGGTAGACTTGAGTCATTGTGAAGTGGTTAGAATGTGTGGTGTAGCGGTGAAATGCATAGATATCACACAGAATACCAATTGCGAAGGCAGATCACTAACAATGTACTGACGCTCATGGACGAAAGCGTGGGGAGCGAACAGG" | 
             rownames(.) == "TGGGGAATATTGCACAATGGGGGAAACCCTGATGCAGCAACGCCGCGTGGAGGATGACACATTTCGGTGCGTAAACTCCTTTTATATAGGAAGATAATGACGGTACTATATGAATAAGCGCCGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGGGCGCAAGCGTTACTCGGAATCACTGGGCGTAAAGAGCGTGTAGGCGGGTTAATAAGTTTGAAGTGAAATCCTATGGCTCAACCATAGAACTGCTTTGAAAACTGTTAACCTAGAATATGGGAGAGGTAGATGGAATTTCTGGTGTAGGGGTAAAATCCGTAGAGATCAGAAGGAATACCGATTGCGAAGGCGATCTACTGGAACATTATTGACGCTGAGACGCGAAAGCGTGGGGAGCAAACAGG" | 
             rownames(.) == "TGGGGAATCTTAGACAATGGGGGAAACCCTGATCTAGCCATGCCGCGTGAGTGACGAAGGCCTTAGGGTCGTAAAGCTCTTTCGCTGGGGAAGATAATGACTGTACCCAGTAAAGAAACCCCGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGACTGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACGGCCTCCAAAACTATCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGCGTGGGGAGCAAACAGG" | 
             rownames(.) == "TGAGGAATATTGGACAATGGGCGCAAGCCTGATCCAGCCATGCCGCGTGCAGGAAGAATGCCCTATGGGTTGTAAACTGCTTTTATTTGGGAATAAACCTCCTTACGTGTAGGGAGCTGAATGTACCAAACGAATAAGCACCGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGGGTGCAAGCGTTATCCGGAATCATTGGGTTTAAAGGGTCCGCAGGCGGGCCTATAAGTCAGTGGTGAAATCCCATCGCTTAACGATGGAACTGCCATTGATACTGTAGGTCTTGAATTCGGTCGAAGTGGGCGGAATGTGTCATGTAGCGGTGAAATGCATAGATATGACACAGAACACCGATAGCGAAGGCAGCTCACTAGGCCTGGATTGACGCTCAGGGACGAAAGCGTGGGGAGCGAACAGG") %>% 
  base::t() %>% 
  as.data.frame() %>% 
  dplyr::rename("Coraliomargarita" = "TTTCGAATCATTCACAATGGGGGAAACCCTGATGGTGCAACGCCGCGTGGGGGATGAAGGCCTTCGGGTTGTAAACCTCTGTCACCAAGGAGCAACAAGCCGGTTCATAGCCGGCCCTGAGTTAACTTGGAGAGGAAGCAGTGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGACTGCAAGCGTTACTCGGATTCACTGGGCGTAAAGGGTGCGTAGGCCGCTAAGCGTGTCGGGTGTGAAATCTCGGGGCTCAACCTCGAAACTGCGCCCGAAACTGTTTAGCTAGAGTGTCGGAGAGGTAAGCGGAATTCCAGGTGTAGCGGTGAAATGCGTAGATATCTGGAGGAACACCAATGGCGAAGGCAGCTTACTGGACGACAACTGACGCTGAGGCACGAAAGCGTGGGTAGCGAAAGGG",  
         "NS4_marine_group" = "TGAGGAATATTGGACAATGGACGAAAGTCTGATCCAGCCATGCCGCGTGCAGGATGACGGCCCTATGGGTTGTAAACTGCTTTTATACAGGAAGAAACACTCCCACGTGTGGGGGCTTGACGGTACTGTACGAATAAGGATCGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGGATCCAAGCGTTATCCGGAATTATTGGGTTTAAAGGGTCCGCAGGCGGTCTATTAAGTCAGAGGTGAAATCTTGCAGCTCAACTGTAAAATTGCCTTTGATACTGGTAGACTTGAGTCATTGTGAAGTGGTTAGAATGTGTGGTGTAGCGGTGAAATGCATAGATATCACACAGAATACCAATTGCGAAGGCAGATCACTAACAATGTACTGACGCTCATGGACGAAAGCGTGGGGAGCGAACAGG",
         "Arcobacter" = "TGGGGAATATTGCACAATGGGGGAAACCCTGATGCAGCAACGCCGCGTGGAGGATGACACATTTCGGTGCGTAAACTCCTTTTATATAGGAAGATAATGACGGTACTATATGAATAAGCGCCGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGGGCGCAAGCGTTACTCGGAATCACTGGGCGTAAAGAGCGTGTAGGCGGGTTAATAAGTTTGAAGTGAAATCCTATGGCTCAACCATAGAACTGCTTTGAAAACTGTTAACCTAGAATATGGGAGAGGTAGATGGAATTTCTGGTGTAGGGGTAAAATCCGTAGAGATCAGAAGGAATACCGATTGCGAAGGCGATCTACTGGAACATTATTGACGCTGAGACGCGAAAGCGTGGGGAGCAAACAGG", 
         "Marivivens" = "TGGGGAATCTTAGACAATGGGGGAAACCCTGATCTAGCCATGCCGCGTGAGTGACGAAGGCCTTAGGGTCGTAAAGCTCTTTCGCTGGGGAAGATAATGACTGTACCCAGTAAAGAAACCCCGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGACTGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACGGCCTCCAAAACTATCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGCGTGGGGAGCAAACAGG", 
         "Salinirepens" = "TGAGGAATATTGGACAATGGGCGCAAGCCTGATCCAGCCATGCCGCGTGCAGGAAGAATGCCCTATGGGTTGTAAACTGCTTTTATTTGGGAATAAACCTCCTTACGTGTAGGGAGCTGAATGTACCAAACGAATAAGCACCGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGGGTGCAAGCGTTATCCGGAATCATTGGGTTTAAAGGGTCCGCAGGCGGGCCTATAAGTCAGTGGTGAAATCCCATCGCTTAACGATGGAACTGCCATTGATACTGTAGGTCTTGAATTCGGTCGAAGTGGGCGGAATGTGTCATGTAGCGGTGAAATGCATAGATATGACACAGAACACCGATAGCGAAGGCAGCTCACTAGGCCTGGATTGACGCTCAGGGACGAAAGCGTGGGGAGCGAACAGG") %>% 
  rownames_to_column(var = "ID")), by = "ID") 
```

### Centre and scale numeric variables using defined function.

```{r, warning=F, message=F,  results='hide'}
centre_and_scale <- function(data1){
# get numeric variables
data2 <- data1 %>% 
  select_if(is.numeric)
# centering and scaling over variables
data3 <- sapply(data2, function(x) scale(x, center=T, scale = 2*sd(x))) %>% 
  as.data.frame() %>% 
  rownames_to_column("RowID")
# join scaled/centred data to non-numeric data
data1 %>% 
  select_if(negate(is.numeric)) %>%
  rownames_to_column("RowID") %>% 
  left_join(data3, by = "RowID") %>%
  select(-RowID)
}

glm_data <- metadata_with_taxa %>% centre_and_scale()
```


### Explore relationship between parasite and variables

```{r, eval=F}
metadata_with_taxa %>% 
  ggplot(aes(x = ddPCR, y = Rainfall)) + 
  geom_point() +
  scale_x_log10() + 
  geom_smooth(method = "lm", se = T) + 
  scale_y_log10()

metadata_with_taxa %>% 
  ggplot(aes(x = ddPCR, y = RDO_Conc_mgL)) + 
  geom_point() +
  geom_smooth(method = "lm", se = T) 

```


### Test for multicollinearity: define the `corvif()`function that takes metadata and creates a linear model to see if any collinearity exists between variables.

```{r}
# myvif
myvif <- function(mod) {
  v <- vcov(mod)
  assign <- attributes(model.matrix(mod))$assign
  if (names(coefficients(mod)[1]) == "(Intercept)") {
    v <- v[-1, -1]
    assign <- assign[-1]
  } else warning("No intercept: vifs may not be sensible.")
  terms <- labels(terms(mod))
  n.terms <- length(terms)
  if (n.terms < 2) stop("The model contains fewer than 2 terms")
  if (length(assign) > dim(v)[1] ) {
    diag(tmp_cor)<-0
    if (any(tmp_cor==1.0)){
      return("Sample size is too small, 100% collinearity is present")
    } else {
      return("Sample size is too small")
    }
  }
  R <- cov2cor(v)
  detR <- det(R)
  result <- matrix(0, n.terms, 3)
  rownames(result) <- terms
  colnames(result) <- c("GVIF", "Df", "GVIF^(1/2Df)")
  for (term in 1:n.terms) {
    subs <- which(assign == term)
    result[term, 1] <- det(as.matrix(R[subs, subs])) * det(as.matrix(R[-subs, -subs])) / detR
    result[term, 2] <- length(subs)
  }
  if (all(result[, 2] == 1)) {
    result <- data.frame(GVIF=result[, 1])
  } else {
    result[, 3] <- result[, 1]^(1/(2 * result[, 2]))
  }
  invisible(result)
}

# corvif
corvif <- function(data) {
  data <- as.data.frame(data)

  form    <- formula(paste("fooy ~ ",paste(strsplit(names(data)," "),collapse = " + ")))
  data  <- data.frame(fooy = 1 + rnorm(nrow(data)) ,data)
  lm_mod  <- lm(form,data) # runs linear model with above formula and metadata
  
  cat("\n\nVariance inflation factors\n\n")
  print(myvif(lm_mod))
}
```

```{r}
# full 
glm_data %>% 
  select(Temperature_C, pH,  
         RDO_Conc_mgL, Rainfall, Salinity_PSU, 
         NS4_marine_group, Salinirepens, Marivivens, 
         Coraliomargarita, Arcobacter) %>% 
  corvif()

# final model
glm_data %>% 
  select(RDO_Conc_mgL, Rainfall, 
         Salinity_PSU, NS4_marine_group,   
         Coraliomargarita, Arcobacter) %>% 
  corvif()
```


## Fit Model.

```{r, warning=F, message=F}
global <- lme4::glmer(parasite_burden ~ RDO_Conc_mgL + 
                      Rainfall + Salinity_PSU + NS4_marine_group + 
                      Coraliomargarita + Arcobacter + (1|Date), 
                      data = glm_data, family = "binomial") 
summary(global)
```

### Goodness of fit and R2.

```{r, warning=F, message=F}
gof(global)
r.squaredGLMM(global)
```


## Backwards selection.

```{r, warning=F, message=F}
dfun(drop1(global))

global2 <- lme4::glmer(parasite_burden ~ RDO_Conc_mgL + 
                      Rainfall + NS4_marine_group + 
                      Coraliomargarita + (1|Date), 
                      data = glm_data, family = "binomial") 

dfun(drop1(global2))

summary(global2) # final model
```

### Goodness of fit and R2.

```{r, warning=F, message=F}
gof(global2)
r.squaredGLMM(global2)
```


### Export summary as a table
```{r, eval=F}
tab_model(global2, show.se = TRUE, string.se = "Standard Error", show.ci = FALSE, 
          show.re.var = FALSE, show.ngroups = FALSE, show.icc = FALSE, 
          title = "Generalised logistic regression model", 
          file = "C:/Users/Jacob/Desktop/Other_Projects/Aquaculture_Microbiome/Outputs/GLM.doc")
```


## Reintgeration to calculate estimates, standard error and p values for each of the variables removed during backwards selection.

```{r, eval=F}
lme4::glmer(parasite_burden ~ RDO_Conc_mgL + Salinity_PSU +
                      Rainfall + NS4_marine_group + 
                      Coraliomargarita + (1|Date), 
                      data = glm_data, family = "binomial") %>% 
  tab_model(show.se = TRUE, string.se = "Standard Error", show.ci = FALSE, 
            show.re.var = FALSE, show.ngroups = FALSE, show.icc = FALSE, 
            title = "Generalised logistic regression model", 
            file = "GLM2.doc")
```

```{r, eval=F}
lme4::glmer(parasite_burden ~ RDO_Conc_mgL + Arcobacter +
                      Rainfall + NS4_marine_group + 
                      Coraliomargarita + (1|Date), 
                      data = glm_data, family = "binomial") %>% 
  tab_model(show.se = TRUE, string.se = "Standard Error", show.ci = FALSE, 
           show.re.var = FALSE, show.ngroups = FALSE, show.icc = FALSE, 
           title = "Generalised logistic regression model", 
           file = "GLM3.doc")
```


## Visualisation.

### Plot effects
```{r}
plot(allEffects(global2))
```


### Calculate the mean and standard deviation for each variable.

```{r, warning=F, message=F}
plot_data <- metadata_with_taxa %>%
  group_by(parasite_burden) %>% 
  select_if(is.numeric) %>% 
  summarise_all(mean) %>% 
  left_join(
    metadata_with_taxa %>%  
      group_by(parasite_burden) %>% 
      select_if(is.numeric) %>% 
      summarise_all(sd) %>% 
      rename_at(2:ncol(.), toupper))
```

### Generate plot.

```{r, warning=F, message=F}
grid.arrange(
(ggplot(plot_data, aes(x = parasite_burden, y = Coraliomargarita)) + 
geom_pointrange(aes(ymin = Coraliomargarita - CORALIOMARGARITA, 
                    ymax = Coraliomargarita + CORALIOMARGARITA)) +
  xlab("") +
  ylab("Coraliomargarita") +
  geom_text(aes(label = "*", y = 4500, x = 1.5, fontface = "bold", size = 20)) +
  theme(legend.position = "none")) %>% 
  annotate_figure(fig.lab = "A", fig.lab.face = "bold", fig.lab.size = 20),

(ggplot(plot_data, aes(x = parasite_burden, y = Arcobacter)) + 
geom_pointrange(aes(ymin = Arcobacter - ARCOBACTER, 
                    ymax = Arcobacter + ARCOBACTER)) +
  xlab("") +
  ylab("Arcobacter") +
  theme(legend.position = "none")) %>% 
  annotate_figure(fig.lab = "B", fig.lab.face = "bold", fig.lab.size = 20),

(ggplot(plot_data, aes(x = parasite_burden, y = NS4_marine_group)) + 
geom_pointrange(aes(ymin = NS4_marine_group - NS4_MARINE_GROUP, 
                    ymax = NS4_marine_group + NS4_MARINE_GROUP)) +
  xlab("") +
  ylab("NS4 marine group") +
  theme(legend.position = "none")) %>% 
  annotate_figure(fig.lab = "C", fig.lab.face = "bold", fig.lab.size = 20),

(ggplot(plot_data, aes(x = parasite_burden, y = Rainfall)) + 
geom_pointrange(aes(ymin = Rainfall - Rainfall, 
                    ymax = Rainfall + Rainfall)) +
  xlab("") +
  ylab("Rain (m)") +
  geom_text(aes(label = "*", y = .1, x = 1.5, fontface = "bold", size = 20)) +
  theme(legend.position = "none")) %>% 
  annotate_figure(fig.lab = "D", fig.lab.face = "bold", fig.lab.size = 20),

(ggplot(plot_data, aes(x = parasite_burden, y = RDO_Conc_mgL)) + 
geom_pointrange(aes(ymin = RDO_Conc_mgL - RDO_CONC_MGL, 
                    ymax = RDO_Conc_mgL + RDO_CONC_MGL)) +
  xlab("") +
  ylab("RDO (mg/L)") +
  theme(legend.position = "none")) %>% 
  annotate_figure(fig.lab = "E", fig.lab.face = "bold", fig.lab.size = 20),

(ggplot(plot_data, aes(x = parasite_burden, y = Salinity_PSU)) + 
geom_pointrange(aes(ymin = Salinity_PSU - SALINITY_PSU, 
                    ymax = Salinity_PSU + SALINITY_PSU)) +
  xlab("") +
  ylab("Salinity (PSU)") +
  theme(legend.position = "none")) %>% 
  annotate_figure(fig.lab = "F", fig.lab.face = "bold", fig.lab.size = 20),

ncol = 3, bottom = textGrob("Cryptocaryon abundance (eDNA)", gp = gpar(fontsize = 12))
)
```

```{r, warning=F, message=F}
grid.arrange(
(ggplot(plot_data, aes(x = parasite_burden, y = Coraliomargarita)) + 
geom_pointrange(aes(ymin = Coraliomargarita - CORALIOMARGARITA, 
                    ymax = Coraliomargarita + CORALIOMARGARITA)) +
  xlab("") +
  ylab("Coraliomargarita") +
  geom_text(aes(label = "*", y = 4500, x = 1.5, fontface = "bold", size = 20)) +
  theme(legend.position = "none")) %>% 
  annotate_figure(fig.lab = "A", fig.lab.face = "bold", fig.lab.size = 20),

(ggplot(plot_data, aes(x = parasite_burden, y = Rainfall)) + 
geom_pointrange(aes(ymin = Rainfall - Rainfall, 
                    ymax = Rainfall + Rainfall)) +
  xlab("") +
  ylab("Rain (m)") +
  geom_text(aes(label = "*", y = .1, x = 1.5, fontface = "bold", size = 20)) +
  theme(legend.position = "none")) %>% 
  annotate_figure(fig.lab = "B", fig.lab.face = "bold", fig.lab.size = 20),

ncol = 2, bottom = textGrob("Cryptocaryon abundance (eDNA)", gp = gpar(fontsize = 12))
)
```

**Finished.**